A method for inference of fundamental stellar parameters using high resolution spectra

### 2014-01-19 paper/paper.tex
Edited half the paper

\section{Introduction}

@@

\subsection{Similarity techniques}
Many of existing techniques were written to analyze low to medium resolution ($R < $ 10,000) spectra. I am calling them similarity techniques because they compare characteristics of spectra, not actual pixels.

\paragraph{Line indexes} By cataloguing the equivalent widths of chosen spectral lines into indexes, one can correlate the index with stellar properties and identify features that are sensitive temperature, surface gravity, or metallicity. Then, one matches the indecies of a new star against these relationships and derives a set of stellar parameters. A recent example using M dwarfs is \citet{nci+14}. This technique works, but there are many things that could be improved. First, the selection of which lines to measure is arduous and based upon expert knowledge of previously catalogued stars. This is both a disadvantage and an advantage. One one hand by choosing a subset of lines to fit, one may be unwittingly throwing away useful information. On the other hand, it allows the modeller to discard regions of the spectrum that do not follow a simple trend with stellar parameters. Second, while measuring equivalent widths is easy to do, this discards any useful information contained in the shape of the line profile.

@@

\subsection{Pixel to pixel techniques using spectral libraries}
Many high quality synthetic spectral grids are now available. Nearly all synthetic stellar libraries are parameterized by the ``fundamental'' stellar parameters of effective temperature, surface gravity, metallicity, and alpha enhancement. We define this combination as 
\begin{equation}
  \vg = \{T_{\rm eff}, \log(g), \Z, \A \}
\end{equation}
Naively, it makes sense to compare an observed spectrum against spectra from the synthetic grid. Using some goodness of fit statistic one can identify the $\vg$ that best reproduce the observed spectrum.

@@

\paragraph{``Bayesian'' methods} do exist. \citet{sdm+07} developed a hierarchical Bayesian approach that purportedly accounts for both statistical and systematic errors in the estimation of stellar parameters. I need to work through their math more carefully, but it seems as though they use a hierarchical model for the scatter between ``true'' and observed spectrum, and the unknown scatter in the relationship between $\vg$ and the $\fM$. They claim to account for systematics, but it looks like these are systematics errors due to the performance of the instrument, such as a bad flux calibration. They use infrared spectra at low resolution ($R \sim $ 1,200), and a weighted $\chi^2$ function for goodness of fit. As far as I can ascertain, the actually pixel-to-pixel fitting does not account for systematic errors in regions of the model spectrum.

@@

\paragraph{SPC} A very popular method behind the Kepler follow-up program is Stellar Parameter Classification (SPC) \citep{blj+12}). This compares high resolution optical echelle spectra ($R \sim$ 45,000) to high resolution synthetic grids computed with Kurucz models by a cross-correlation method. The best fit parameters and uncertainties are calculated by the relative values of the cross-correlation coefficient. It is uncertain how statistically justified this uncertainty estimate is. Currently SPC uses only a small region of the optical spectrum ($5000 - 5300$\AA) where the spectral library has been carefully tweaked\footnote{changes to line centers and $\log(gf)$ values were on the table} to match high resolution models of the Sun and Vega.

\paragraph{General trends}
Successful pixel-to-pixel methods either mask bad spectral lines or limit themselves to a narrow region of the spectrum, where they can be relatively certain that the spectral library does not contain gross systematical errors. This implicit masking brings up similar concerns to those of \citet{mga13}: how do you know that you are fitting the region of the spectrum that contains the most information about the stellar properties?

\subsection{Line fitting using spectral synthesis}
This is an improvement over the line index method, since now one includes the information contained in the line profile. With quality stellar spectral models, one can identify a few well-modelled lines to fit. By choosing a few good lines to fit, one could determine stellar parameters by matching line profiles for a few select lines \citep{dj03}. However, this is still subject to the need to identify promising regions of the spectrum.

MOOG\footnote{\url{http://www.as.utexas.edu/~chris/moog.html}} \citep{sne73}, SME\footnote{\url{http://www.stsci.edu/~valenti/sme.html}} \citep{vp96}, actually synthesize line profiles on the fly, compared to using a library of spectral models. This has the advantage of enabling individual tweaks to elemental abundances. And interpolation in the model atmosphere space, rather than the spectral line space. However, for many young T-Tauri stars, the spectrum is contaminated by accretion, veiling, and fractional star spot coverage, which can be a nuisance and cause systematic errors if traditional techniques are used. Matching equivalent widths. Interestingly, MOOG sythesizes a spectrum on the fly.

SME allows adjustment of oscillator strengths. We could download the synthesis code for SME and synthesize our own spectra on the fly. The biggest impediment here would be that the simplest atmospheres will not get certain lines correct. The powerful synthesis codes (PHOENIX) will get the atmospheres and lines correct, but they take a long time to run (~hours) per model, and the code is not publicly available. To do a full-spectrum fit for these sorts of codes would likely be too computationally difficult. However, it would allow access to fitting in the parameter space that we want.

\subsection{Stellar parameter estimation in a utopia}

Although the spectral synthesis domain is the correct domain to solve for stellar parameters, the radiation transfer is too slow.

The fitting difficulties of the previous section are a result of trying to fit data to a model in a smaller parameter space. Yes, the combination of $\vg$ is an excellent proxy to describe a star, but in truth (and at high resolution) each star is slightly different. If we think back to first principles, the ``model'' that is actually producing the synthetic spectrum has many, many parameters that have been collapsed down into $\vg$. These parameters include the atmospheric structure (relations of temperature and pressure as a function of stellar depth), individual elemental abundances (all models quote some value relative to solar, a fraction of models allow changes in alpha abundances), and atomic constants describing the opacity contribution of a line. \emph{All} of these parameters have uncertainties in them, some of these are rather dramatic (one continually hears lamentation about the opacities and line lists\dots), yet for any given stellar model these parameters are \emph{fixed}. There may yet be unconsidered stellar physics necessary to explain a small fraction of lines (say for example the cores of Ca lines), but the field of stellar modelling has progressed to the state that I would have the confidence to say that if a group focused all of their attention on a single line, and using only their current physical models and making tweaks to the afore-mentioned lines, they could produce a satisfactory model.

In reality, this has been done using the Kurucz models in an unpublished grid of stellar spectra, over the range $5000 - 5300$ \AA.

The real problem is much more rich, as is the amount of information one could extract from a high resolution optical spectrum. One would want to fit for individual abundances, marginalize over oscillator strengths.

The unfortunate reality is that sampling in the stellar model space for the full optical spectrum is not yet a computational reality. There do exist codes (MOOG) that can do this approach but only for a small ($\sim$ a few lines) at once, and using simplifications. Through collaboration with a stellar atmospheric expert, such an approach may soon be a possibility. If the computation of a given model could be reduced to less than a minute, efficient MCMC sampling parallelized on a cluster (using \texttt{emcee}, for example). However, as it stands without a collaborator, I do not have the expertise nor the time to pursue such an approach. Therefore, we are reduced to fitting in the space of $\vg$ and devising ways to deal with the issues.

@@

Pixel-to-pixel approaches have the potential to be much more powerful than other historical stellar parameter estimation techniques. When comparing directly in the data space, we have the ability to add in other physics as we see fit.

because we can fit multiple stars at once (say for unresolved binary or multiple systems), fit for velocity components, etc. Many of the most interesting systems for which we would want to determine stellar parameters are in multiple systems anyway, since we can dynamically determine masses for these systems and thus compare to evolutionary models. Determine veiling. And a perhaps wooden-headed belief that using the full spectrum \emph{should} give us a more accurate and less-biased posterior estimate. Such a belief may be foolish and in reality different parts of the spectrum could give completely different parameter estimates.

Many, many methods exist for comparison of stellar spectra, including many ``Bayesian'' approaches in the pixel-to-pixel space. The nature of the Bayesian-ness, however, is usually limited to priors on temperature, surface gravity, and metallicity. These approaches do not explicitly state the likelihood function that they use, or if they do, it is a $\chi^2$ function.

The more successful of these models use low or moderate resolution spectra ($R \sim$ 1,000 to 10,000 ish). Or, a grid over a narrow range of wavelength space with finely tuned atomic constants.

\textbf{The true fundamental problem with this approach is that the stellar spectra are not a perfect representation of reality}. Assumptions about Poisson errors break down, and fits become contaminated by systematics. There are types of errors which occur.

FIGURE: 4 Panel Figure, from left to right, showing Class 0 through III errors. Use Mg b lines, other big line near Mg b triplet, order 24 line, and then Fe I/II line complex that doesn't fit properly.

@@

\paragraph{Class II} A spectral line appears in the model but not in the data over a wide range of probable stellar parameters (1000 K in $T_{\rm eff}$, and 1 dex in $\log(g)$, $\Z$, and $\A$. This is a more puzzling error, because presumably atomic constants would be much more likely to be \emph{missing} than added in by accident. Presumably this is a product of bad atmospheric structure (temperature or pressure is wrong at some point?) that excites a line when it shouldn't otherwise.

\paragraph{Traditional ways to model Class I and Class II errors} are a form of sigma clipping. Mann et al. compares the stars to models over a wide range of parameters. One can track how these lines might be bad because Class I residuals (data - model)/sigma will always be negative (since we are dealing with absorption lines only) and not change significantly with $\vg$, since we the model can only produce continuum. Class II residuals will always be positive and should decrease with decreasing metallicity, since lower metallicity will decrease the strength of all lines.

If a Class I or II error occurs on a line, that region of the spectrum is relatively useless for determining spectral parameters. Authors like Mann et al. or Kholeva et al. find these regions and simply mask them out of the final computation. Kholeva et al., and others approach in an iterative manner. Kholeva et al. approaches are more sophisticated because they track convergence. This is a sigma clipping approach, and while we could go and model these a la Hogg 2010, the approach is not super justified, result is probably not too wrong (similar to actual sigma clipping). Given that Kholeva et al. track convergence maps, I would place a higher degree of confidence in the answer.

In our later approaches in Sec XX, how might we put a prior on $\Delta \vg$, the range of parameters over which the residuals will be correlated? Maybe we could Gibbs sample in the same way as the bad flags. We'd have to keep a bunch of spectra with varying $\vg$ in memory to make this comparison easy.

\paragraph{Class III} In truth, this is just a superset of both Class I and Class II. Class III errors are when the line strengths between the data and model do not match at a $\vg$ that fits a majority of lines. What makes Class III errors truly pernicious is that for a specific line, there does exist a set of $\vg$ that could properly fit the line, however this parameter combination would result in bad fits for the majority of the other lines. Class I and II errors are easy to identify since the lines will always be bad across a wide range of $\vg$.

In addition, it is difficult to determine the ensemble of lines that suffer from Class III errors, since the strength of some lines will be under-predicted while others will be over-predicted. How do we know which lines to trust? If we take an approach similar to fixing Class I and II errors, and mask out bad regions, it would be very difficult to distinguish Class III errors from \emph{truthful} lines at an incorrect $\vg$. For example, if one imagines using a cleaning algorithm as described for Class I and Class II errors, the identification of Class III errors will be sensitive to the starting guesses of $\vg$.

Because these Class III lines can still produce a significant residual spike, they tend to distort and in some cases artificially peak the posterior distribution. There is no one parameter combination that will fit all lines perfectly. Certainly there is a parameter combination that will produce a reasonable fit to 90\% of the lines to within 3 sigma residuals, but there will still be a fair number of both ``high'' and ``low'' regions. These lines need to be identified and dealt with, but how?

How then, should the best fit be determined? Surely some of the remaining lines are ``bad.'' But in this regime, determination of which lines are bad will bias the results in a certain direction.

Also, which lines should one actually fit? The stronger lines? Did the modelers themselves give more attention to the atomic constants which produce the strong lines?

To summarize, the fundamental problem with comparing high resolution optical stellar spectra to grids of synthetic stellar spectra is that no one model $\vg$ will fit all of the lines. Deciding how to model the lines is then the difficult part. Different groups have sigma-clipped lines that are obviously bad, and this does improve the procedure dramatically. In such a case, the procedure of sigma-clipping doesn't gall with me too badly. However, the more difficult problem is dealing with Class III errors properly, something that no modeler has mentioned or successfully implemented in a statistically justified manner.

In a spectrum fit, many lines do reasonably fit within the Poisson errors. For consistency, we will denote these as ``Class 0'' lines, where there is no error.

\subsection{Testing the accuracy of stellar libraries}
Gaia team members are panicking about the upcoming flood of data.

\citet{mc07} compare many spectral grids to determine their weaknesses. \citep{svt+11} discuss the range of stellar grids available for Gaia.

\citep{cbm+05} is the Coehlo grid.

\section{Instrumental effects and ``nuisance'' parameters in the limit of a perfect stellar model}

The parameter vector is comprised of both stellar parameters and calibration parameters $\vt = \{\vstar, \vN \}$. Stellar parameters include $\vstar = \{T_{\rm eff}, \log(g), [{\rm Fe}/{\rm H}], v \sin i, v_z, A_v, R_\star^2/d^2 \}$ while the data vector $\vD = \fD(\lambda)$ represents a high resolution spectrum. For demonstration purposes in this paper, we use a 51-order echelle spectrum from the \emph{TRES} spectrograph at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$), which spans the full optical range ($3800-9000$\AA), though our technique is applicable to any high resolution optical/near-IR spectrum.

FIGURE showing the high resolution, raw $R =$ 500,000 spectrum compared to the one convolved with a kernel down to $R =$ 50,000. This shows that lines are blended, but it is important to keep in mind that we do have access to the high resolution models when talking about bad pixels or bad line-species. For example, we could inject bad lines in this space.

\subsection{Processing TRES Data}
We flux calibrate TRES data, but because the reduction pipeline does not produce a ``sigma spectrum,'' we instead use the Poisson errors on each pixel. By taking the raw, still-blazed spectrum, in units of ``counts,'' we can find the Signal-to-Noise ratio by,
\begin{equation}
  S/N = \sqrt{ {\rm cts}}
\end{equation}
then the Noise-to-Signal ratio is the inverse of this
\begin{equation}
  N/S = \frac{1}{\sqrt{ {\rm cts}}}
\end{equation}
then the errors on the flux-calibrated spectrum (in ergs/s/A/cm2) is simply the flux spectrum multiplied by N/S.

\subsection{Applied to a Gaussian likelihood function}

Although per-pixel photon counting errors are Poisson, we can safely assume that we have enough counts that we are in the Gaussian limit, although see the discussion in \S\ref{sec:residuals}. Then our likelihood function becomes a pixel-by-pixel $\chi^2$ comparison between the data spectrum $\fD$, and the model spectrum $\fM$, summed over the wavelength axis. 
\begin{equation}
 p(\vD | \vt) = {\cal L} \propto \exp \left(-\frac{\chi^2}{2} \right)
\end{equation}
\begin{equation}
  \chi^2 = \sum_\lambda \left [\frac{ \fD(\lambda) - k(\lambda | \vN) \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
\end{equation}
With flat priors on $\vstar$ and $\vN$ (an assumption we will later relax) the logarithm of our posterior probability function is then
\begin{equation}
  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{ \fD(\lambda) - k(\lambda | \vN) \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
  \label{eqn:lnprob}
\end{equation}
To generate a model spectrum given a set of parameters $\fM(\vt)$, a model spectrum is created by linearly interpolating in $T_{\rm eff}, \log(g), [{\rm Fe}/{\rm H}]$ space between grid points of instrumentally broadened, oversampled PHOENIX model stellar spectra, rotationally broadened using Fourier techniques, Doppler shifted, extinction corrected, and downsampled using spline interpolation to the exact pixels of the TRES spectrum.

$\sigma(\lambda)$ is the Poisson photon counting errors. For some data reduction pipelines, a per-pixel ``sigma spectrum'' that accounts for spectral extraction errors due to night sky line contamination or low signal to noise is available, and should be used instead. For more discussion of the noise sources beyond statistical Poisson errors, see \S\ref{sec:residuals}.

>>>>

\section{Introduction and review of existing methods}

@@

\subsection{Classification techniques}
Many of existing techniques were written to analyze low to medium resolution ($R < $ 10,000) spectra. I am calling them similarity techniques because they focus on determining and recognizing patterns of spectra useful in classification.

\paragraph{Line indexes} By cataloguing the equivalent widths of chosen spectral lines into indexes, one can correlate the index with stellar properties and identify features that are sensitive temperature, surface gravity, or metallicity. Then, one matches the indecies of a new star against these relationships and derives a set of stellar parameters. A recent example using M dwarfs is \citet{nci+14}. This technique works, but there are many things that could be improved. First, the selection of which lines to measure is arduous and based upon expert knowledge of previously catalogued stars. This is both a disadvantage and an advantage. On one hand, by choosing a subset of lines to fit, one may be unwittingly throwing away useful information. On the other hand, it allows the modeller to discard regions of the spectrum that do not follow a simple trend with stellar parameters. Second, while measuring equivalent widths is easy to do, this discards any useful information contained in the shape of the line profile.

@@

\paragraph{Machine learning}
Many machine learning techniques exist and classify spectra by comparing patterns between observed spectra and a training set of spectra (which could be real spectra of stars classified by an expert or a synthetic library that has known $\vg$). However, my first impression is that these tools are designed to handle large collections of spectra from stellar surveys such as SDSS \citep{sam+10}. In our immediate project involving T Tauri stars, we only have a few ($\lesssim 20$) stars that we wish to study in exquisite detail, incorporating specific parameters such as $v \sin i$, extinction, and veiling. However, I am a novice when it comes to machine learning so I may be missing something here.

\subsection{Pixel to pixel techniques using synthetic spectral libraries}
\label{sec:pix}
Many high quality synthetic spectral grids are now available. Nearly all synthetic stellar libraries are parameterized by the ``fundamental'' stellar parameters of effective temperature, surface gravity, metallicity, and alpha enhancement. We define this combination as 
\begin{equation}
  \vg = \{T_{\rm eff}, \log(g), \Z, \A \}
\end{equation}
Naively, it makes sense to compare an observed spectrum against spectra from the synthetic grid. Using some goodness of fit statistic one can identify the $\vg$ that best reproduce the observed spectrum.

@@

\paragraph{``Bayesian'' methods} do exist. \citet{sdm+07} developed a hierarchical Bayesian approach to compare infrared spectra at low resolution ($R \sim $ 1,200) to a synthetic grid in order to determine stellar parameters $\vg$. Their model is hierarchical in that it has parameters describing the statistical and systematic uncertainties of the observed spectrum. The systematic errors are due to the imperfect performance of the instrument, for example a bad flux calibration. They use an ``empirical'' Bayesian model for the noise, meaning that they use data to estimate the prior parameters for the noise \citep{gcs+04}. They use a $\chi^2$ likelihood function and allow for ``shrinkage'' of their hierarchical parameters if there is good agreement between observed and synthetic spectra. Disappointingly, they do explicitly account for systematic errors in the synthesis of actual lines, for example due to wrong atomic constants. One could imagine having a systematic error parameter which describes the scatter in the relationship between $\fM(\lambda | \vg)$ and the ``true'' physical spectrum in the absence of noise $\fD$. Nor do they account for any systematic effect that likely exists between $\vg$ and the true physical stellar parameters (as far as $T_{\rm eff}$ has any physical significance) and is a function of which synthetic grid they use\footnote{For example, in our tests with Kurucz vs. PHOENIX spectra we often find a systematic offset in temperature of $\gtrsim 200$ K.}. However, they do account for interpolation errors by testing interpolating spectra in the pixel space and atmospheric interpolations. Then, they select the best model using squared error loss function and a penalty measure for model complexity.

@@

\paragraph{SPC} A workhorse method behind the Kepler follow-up program is Stellar Parameter Classification (SPC) \citep{blj+12}). This compares high resolution optical echelle spectra ($R \sim$ 45,000) to high resolution synthetic grids computed with Kurucz models by a cross-correlation method. The best fit parameters and uncertainties are calculated by the relative values of the cross-correlation coefficient. It is uncertain how statistically justified this uncertainty estimate is. Currently SPC uses only a small region of the optical spectrum ($5000 - 5300$\AA) where the spectral library has been carefully tweaked\footnote{changes to line centers and $\log(gf)$ values} to match high resolution models of the Sun and Vega.

\paragraph{Other methods} \citet{dj03} do not use a Bayesian model, but they do $\chi^2$ fitting to line profiles in the infrared. \citet{bai10} use a forward modelling approach that interpolates a synthetic grid to estimate stellar parameters from low resolution Gaia spectra.

\paragraph{Comparison of synthetic stellar libraries}
Gaia team members are in a mild panic about the upcoming flood of data. \citet{mc07} compare many spectral grids to determine their weaknesses. \citep{svt+11} discuss the range of stellar grids available for Gaia. \citep{cbm+05} is the Coehlo grid. Munari et al 2005, \citep{hwd+13}.

\paragraph{General trends}
Successful pixel-to-pixel methods either mask bad spectral lines or limit themselves to a narrow region of the spectrum, where they can be relatively certain that the spectral library does not contain gross systematical errors (and effectively masking the rest of the spectrum). Cherry picking regions of the spectrum like this evokes similar concerns to those of \citet{mga13}: how do you know that you are fitting the region(s) of the spectrum that deliver the right balance of information about the stellar properties?

\subsection{Line fitting using spectral synthesis}
Direct synthesis methods take a step closer towards stellar parameter estimation in the correct domain. For as large a region of the spectrum as is computationally tractable, these codes synthesize a spectrum for direct comparison with the data. Because this is an expensive operation, one must necessarily choose only a few good lines to fit.

\textbf{MOOG}\footnote{\url{http://www.as.utexas.edu/~chris/moog.html}} \citep{sne73} and \textbf{SME}\footnote{\url{http://www.stsci.edu/~valenti/sme.html}} \citep{vp96} are two popular synthesis codes. Given a set of stellar parameters, elemental abundances, and oscillator strengths, the program will interpolate a model atmosphere and then synthesize a spectrum (over a small region of wavelength). Then, they use a Levenberg-Marquartd algorithm to converge to the best-fit parameters. Their approach has the benefit of enabling individual tweaks to elemental abundances and oscillator strengths, something that is impossible in the aforementioned approaches.

In order to make the direct synthesis approach computationally tractable, one must use the simplest atmospheric models and radiative transfer. Powerful codes like PHOENIX \citep{hwd+13} \emph{should} have more accurate atmospheres and physics, but they take a long time to run ($\sim$ hours) per model, and the code is not publicly available.

\subsection{Stellar parameter estimation in a utopia}
The difficulties encountered by the similarity techniques and synthetic grid comparisons are a result of fitting in the wrong parameter space. Yes, the combination of $\vg$ is an excellent proxy to describe a star, but in truth (and at high resolution) each star is slightly different. The ``model'' that is actually producing the synthetic spectrum has many, many parameters that have been collapsed down into $\vg$. These parameters include the atmospheric structure (relations of temperature and pressure as a function of stellar depth), individual elemental abundances (all models quote a value relative to solar ($\Z$) and some synthetic grids include alpha enhancements ($\A$)), and atomic constants describing the opacity contribution of a line. \emph{All} of these parameters have uncertainties in them, some of them quite dramatic (opacities and line lists are the frequently cited), yet for a synthetic grid these parameters are \emph{fixed}. There may still be some missing physics that would be needed to produce a perfect spectrum within the noise (say for example the cores of Ca lines, or dust settling at the latest spectral types), but my naive impression is that the field of stellar spectral synthesis is mature enough that nearly all lines could be properly reproduced given sufficient scrutiny to the atmosphere and atomic constants.

If we had infinite computational power, a tantalizing approach would be to use a model with stellar parameters $\vg$, atmospheric parameters (such as pressure and temperature profiles), and uncertain atomic constants such as oscillator strengths. One could fit for individual elemental abundances and marginalize over the atomic constants. This would extract a tremendous amount of information from a high resolution spectrum. Unfortunately, the relationship between $\vt$ and the spectrum $\fD(\lambda | \vt)$ is non-linear, meaning that a new spectrum would need to be synthesized for each new $\vt$. It may be possible to make an approximate a linear relationship between say an oscillator strength and the line depth as long as the tweak is minor. If no reasonable approximation can be found, then the radiative transfer becomes an insurmountable bottleneck.

If the approximations could reduce the computation of synthetic spectrum, efficient MCMC sampling parallelized on a cluster might put this utopian vision within reach in the near future. However, without a collaborator with expert knowledge of spectral synthesis, I do not have the expertise to take such an approach. Many high quality synthetic spectral grids do exist, and so our current efforts are focused on addressing the shortfalls of the previously mentioned pixel-to-pixel methods.

@@

Forward modelling approaches (pixel-to-pixel) have the potential to be more powerful than other stellar parameter estimation techniques. When we model directly to the data space, we have the ability to account for instrumental and systematic effects, allowing us to derive a robust estimate of our uncertainty by marginalizing over the nuisance parameters. We can also derive more complicated models to fit unresolved binary or multiple star systems. In fact, many of the most interesting systems for which we might want to determine stellar parameters are in multiple systems since we can dynamically determine masses for these systems and thus compare to evolutionary models.

As discussed in \S\ref{sec:pix}, many forward modelling approaches exist, yet nearly all use a $\chi^2$ likelihood function and most employ some sort of masking scheme. The more successful pixel-to-pixel methods use low or moderate resolution spectra ($R \sim$ 1,000 to 10,000) or a grid over a narrow range of wavelength space with finely tuned atomic constants, eschewing difficult decisions about what to mask. \textbf{The true fundamental problem with a pixel-to-pixel comparison is that the stellar spectra are not a perfect representation of reality}. This may be an obvious statement, but systematic errors in the synthetic spectra means that the previously valid assumption of Poisson errors and a $\chi^2$ likelihood function breaks down. In order to derive unbiased parameter estimates, systematic errors must be identified and properly handled.

\begin{figure}[!htb]
\begin{center}
\includegraphics{lineclasses}
\caption{The four classes of line behavior, from left to right: \textbf{Class 0}: the model behaves as expected. \textbf{Class I}: a spectral line is present in the data but missing from the model (in this case, the line is also blended). \textbf{Class II}: a spectral line exists in the model where there is only continuum in the data. \textbf{Class III}: lines exist in both the data and model, however they are of the wrong strength. Some lines are too strong while others are too weak. The lower panels show the residual spectra in units of the observational uncertainty (Poisson uncertainty only).}
\label{fig:lineclasses}
\end{center}
\end{figure}

\subsection{Classes of synthetic spectral line behavior}
From our forward modelling experiments using the modified Kurucz grid used in \citet{blj+12} and the recent PHOENIX grid \citep{hwd+13}, we have identified ways that synthetic spectra diverge from the our observed data. To start, we will describe these classes of behaviour as being tied to specific spectral lines present against a well-defined continuum. Later on, we will use the classes to describe features more generally to cover the behavior of molecular features in later type spectra. See Figure~\ref{fig:lineclasses} for visual examples of each class of behavior.

\paragraph{Class 0} These lines actually produce a good fit with a residual nearly consistent with the Poisson counting uncertainty (see \S\ref{sec:TRES} for how the data is processed). If every region of the spectrum behaved like this, we might be justified in using a $\chi^2$ likelihood function and calling it a day.

@@

\paragraph{Class II} A spectral line appears in the model but not in the data over a wide range of probable stellar parameters (1000 K in $T_{\rm eff}$, and 1 dex in $\log(g)$, $\Z$, and $\A$. This is a more puzzling error, because presumably atomic constants would be much more likely to be \emph{missing} than added in by accident. Presumably this might be a combination of incorrect atomic data coupled with incorrect atmospheric structure that incorrectly excites a line.

The traditional way to account for Class I and Class II errors is to sigma clip the bad lines in the spectrum \citep{kpb+09, mga13}. A region of the spectrum with Class I or II errors is useless for determining spectral parameters. While sigma-clipping routines are to be avoided, if they can correctly mask these regions, then at least they are moving in the right direction.

There is additional information available to help determine which regions of the spectrum suffer from Class I or II errors.For a normal star with absorption lines against a continuum, Class I residuals will always be negative and not change significantly with $\vg$, since we the model can only produce continuum in that region. Class II residuals will always be positive and their magnitude should decrease with decreasing metallicity, since lower metallicity will decrease the strength of all lines. We could easily store the partial derivatives of an individual synthetic spectrum with respect to $\vg$
\begin{equation}
  \bigl \{ \frac{\partial \fM(\lambda | \vg)}{ \partial T_{\rm eff}}, \frac{\partial \fM(\lambda | \vg)}{ \partial \log(g)}, \frac{\partial \fM(\lambda | \vg)}{ \partial \Z}, \frac{\partial \fM(\lambda | \vg)}{ \partial \A} \bigr \}
  \label{eqn:derivative}
\end{equation}
and coupled with a prior on the range of parameters over which the residuals will be correlated, we could shore up the classification of Class I and II errors.

\paragraph{Class III} errors occur when the line strengths between the data and model do not match at a $\vg$ that fits a majority of the other lines. What makes Class III errors truly pernicious is that for a specific line, there does exist a set of $\vg$ that could properly fit the line, though this parameter combination would result in bad fits for the majority of the other lines. Class I and II errors are easy to identify since the lines will always be bad across a wide range of $\vg$ and have predictable behaviors with change in $\vg$. Class III errors, however, come in both directions: the line can be too strong or too weak. Identifying Class III errors requires making a choice about which lines to trust and how much.

Most of the Class III errors can probably be identified by finding the regions of the spectrum that have the highest residuals after dealing with Class I and II errors, although some mildly incorrect Class III lines will be difficult to distinguish from correct (Class 0) lines that are simply fit with a less probable $\vg$. Once identified, properly handling Class III errors requires careful thought because these regions of the spectrum \emph{still contain information about $\vg$}. If we take an masking approach as with Class I and II errors, we will be throwing it away. It might even be the case that Class III regions contain \emph{the most} information about $\vg$, since these ``delicate regions'' are the most sensitive to changes in $\vg$ at that particular $\vg$ and thus are tricky to tune correctly. In other words, Class III regions might also be regions where the spectrum derivative (Eqn~\ref{eqn:derivative}) is largest.

To summarize, the fundamental problem with comparing high resolution stellar spectra to grids of synthetic stellar spectra is that no one model $\vg$ will fit all of the lines properly. One can either limit their fits to regions of the spectrum that they know their spectral libraries to be accurate (implicitly masking the rest and potentially discarding a large amount of information) or suck it up and forward model the regions where the models perform poorly.

\section{Instrumental effects and ``nuisance'' parameters in the limit of a perfect stellar model}
\label{sec:perfect}
While this document aims to be general in scope, any forward-modelling approach must also discuss the characteristics of the instrument that performed the observations. We are currently modelling data obtained with the \emph{TRES}\footnote{\url{http://tdc-www.harvard.edu/instruments/tres/}} echelle spectrograph, and so we will describe the characteristics of the instrument and the calibration approaches we have taken, although we note that our approach is easily generalized for data obtained with other spectrographs.

\emph{TRES} is a 51-order echelle spectrograph on the 1.5-m Tillinghast reflector at Whipple Observatory. It spans the full optical range ($3800-9000$\AA) at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$). Echelle spectrographs impart a strong blaze function upon each order, where transmission is largest in the center of the order and declines rapidly as one moves towards the edge of the order. The \emph{TRES} pipeline extracts orders such that they span a wavelength region until the transmission reaches $\lesssim$ 25\% of the central transmission in that order. Each order spans $\sim 100$\AA\ and the blue orders overlap while there are gaps between the wavelength coverage of the red orders. This spectra are blaze-corrected by dividing by a blaze function determined from fitting nightly exposures of flat lamps. Then, the spectrum is flux calibrated using sensitivity functions determined from nightly observations of spectrophotometric standards (O stars). The flux-calibrated spectrum is in units of ergs/s/${\rm cm}^2$/\AA.

\paragraph{Processing TRES Data}
\label{sec:TRES}
Because the \emph{TRES} reduction pipeline does not produce a ``sigma spectrum'' like that determined by a standard IRAF reduction, we instead calculate the Poisson uncertainty on each pixel. We can find the signal-to-noise ratio of each pixel using the raw (still-blazed) spectrum, which is in units of ``counts.''
\begin{equation}
  [S/N](\lambda) = \sqrt{ {\rm cts}(\lambda)}
\end{equation}
the Noise-to-Signal ratio is the inverse 
\begin{equation}
  [N/S](\lambda) = \frac{1}{\sqrt{ {\rm cts}(\lambda)}}
\end{equation}
To determine the uncertainty of the flux-calibrated spectrum, we multiply by N/S. The $[N/S](\lambda)$ ``spectrum'' is larger at the edges of each order due to the reduced transmission. Please note that statistic pixel uncertainties derived in this manner neglect regions of the spectrum that might have a higher variance due to sitting atop a night sky line, but short of implementing our own reduction pipeline for \emph{TRES}, there isn't much more we can do at the moment.

\paragraph{Generating a model spectrum} We parameterize the fit to a star by $\vt = \{\vstar, \vN \}$, where stellar parameters include $\vstar = \{T_{\rm eff}, \log(g), \Z, \A, v \sin i, v_z, A_v, R_\star^2/d^2 \}$. The data vector $\vD = \fD(\lambda)$ represents the \emph{TRES} spectrum. To generate a model spectrum given a set of parameters $\fM(\vt)$, a model spectrum is created by first linearly interpolating in $T_{\rm eff}, \log(g), \Z, \A$ space between grid points of instrumentally broadened and oversampled synthetic stellar spectra. Then, the spectrum is rotationally broadened for $v \sin i$ using Fourier techniques, Doppler shifted by $v_z$, extinction corrected for $A_V$, multiplied by $R_\star^2/d^2$ for a solid angle correction, and finally downsampled to the exact pixels of the TRES spectrum using spline interpolation\footnote{Because we are dealing with flux-calibrated data and synthetic spectra, we \emph{resample} to lower resolution. We do not \emph{rebin} because we are dealing with flux-density, not photon counts}.

FIGURE: showing the high resolution, raw $R =$ 500,000 spectrum compared to the one convolved with a kernel down to $R =$ 50,000. This shows that lines are blended, but it is important to keep in mind that we do have access to the high resolution models when talking about bad pixels or bad line-species. For example, we could inject bad lines in this high resolution space.

\subsection{Applied to a Gaussian likelihood function}
For the rest of \S\ref{sec:perfect}, we proceed as if there were no Class I, II, or III errors present in our model so that we may illustrate our technique for dealing with uncertainties in the flux-calibration. Although per-pixel photon counting errors are Poisson, we can safely assume that we have enough counts that we are in the Gaussian limit, although see the discussion in \S\ref{sec:residuals}. Then our likelihood function becomes a pixel-by-pixel $\chi^2$ comparison between the data spectrum $\fD$, and the model spectrum $\fM$, summed over the wavelength axis. 
\begin{equation}
 p(\vD | \vt) = {\cal L} \propto \exp \left(-\frac{\chi^2}{2} \right)
\end{equation}
\begin{equation}
  \chi^2 = \sum_\lambda \left [\frac{ \fD(\lambda) - k(\lambda | \vN) \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
\end{equation}
With flat priors on $\vstar$ and $\vN$ (an assumption we will later relax) the logarithm of our posterior probability function is then
\begin{equation}
  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{ \fD(\lambda) - k(\lambda | \vN) \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
  \label{eqn:lnprob}
\end{equation}
